---
  title: "MedGenome Sample Analysis"
  author: "Biao Li"
  date: "2024-12-02"
  output: 
    html_document:
      toc: true
      # toc_float: true
---

```{r setup, include=FALSE}
library(dplyr)
library(openxlsx)
library(stringr)
```

## Load MedGenome Data Queried from cbioportal

Load the data queried from cbioportal for samples having both WES and RNA-Seq available.

```{r load-medgenome-data}
df_c <- read.csv("./MedGenome_metadata_sent/20241121_medgenome_samples_list.csv", header = TRUE)
```

## Load and Check Data Sent from JJ

### 6236 Dataset (N = 88)

Load the 6236 dataset and clean the `Sample_ID` column:

```{r load-6236}
df_6236 <- read.xlsx("./MedGenome_metadata_from_JJ/RMC-6236 MCT TV% change (cancer discovery).xlsx") %>%
  mutate(Sample_ID = str_remove_all(Model.Name, "-")) %>% # Remove '-' in Sample_ID
  mutate(Sample_ID = str_to_upper(Sample_ID)) %>%           # Convert to uppercase
  mutate(Sample_ID = str_trim(Sample_ID)) %>%               # Trim trailing whitespaces
  mutate(Sample_ID = gsub("LUC$", "", Sample_ID))       # Rename 'LUC$' to ''

# JJ's 6236 samples not found in df_c
df_6236$Sample_ID[!df_6236$Sample_ID %in% df_c$Sample_ID]
```

**Note:** Likely missing samples due to absence of WES or RNA-Seq data.

### 6291 Dataset (N = 40)

Load the 6291 dataset and clean the `Sample_ID` column:

```{r load-6291}
df_6291 <- read.xlsx("./MedGenome_metadata_from_JJ/RMC-6291.xlsx") %>%
  mutate(Sample_ID = str_remove_all(Models, "-")) %>% # Remove '-' in Sample_ID
  mutate(Sample_ID = str_remove_all(Sample_ID, "\\*")) %>% # Remove '*' in Sample_ID
  mutate(Sample_ID = str_to_upper(Sample_ID)) %>%       # Convert to uppercase
  mutate(Sample_ID = str_trim(Sample_ID))               # Trim trailing whitespaces

# JJ's 6291 samples not found in df_c
df_6291$Sample_ID[!df_6291$Sample_ID %in% df_c$Sample_ID]
```

**Note:** Missing sample `ST1384`, likely a typo (either `ST1384` or `ST1348`?).

### 9805 Dataset (N = 30)

Load the 9805 dataset and clean the `Sample_ID` column:

```{r load-9805}
df_9805 <- read.xlsx("./MedGenome_metadata_from_JJ/RMC-9805.xlsx") %>%
  mutate(Sample_ID = str_remove_all(Model, "-")) %>% # Remove '-' in Sample_ID
  mutate(Sample_ID = str_to_upper(Sample_ID)) %>%      # Convert to uppercase
  mutate(Sample_ID = str_trim(Sample_ID))              # Trim trailing whitespaces

# JJ's 9805 samples not found in df_c
df_9805$Sample_ID[!df_9805$Sample_ID %in% df_c$Sample_ID]
```

**Note:** Missing sample `LXFA2889` as in 6236 data.

## Backtrack Samples Contained in Data Queried from cbioportal but Not Found in Data sent from JJ

### 6236 Dataset

Filter samples from `df_c` relevant to the 6236 study and compare with `df_6236`:

```{r backtrack-6236}
df_c_6236 <- df_c %>% 
  filter(!is.na(M_Recist_6236))

table(df_c_6236$Sample_ID %in% df_6236$Sample_ID)
```

### 6291 Dataset

Filter samples from `df_c` relevant to the 6291 study and compare with `df_6291`:

```{r backtrack-6291}
df_c_6291 <- df_c %>% 
  filter(!is.na(M_Recist_6291))

table(df_c_6291$Sample_ID %in% df_6291$Sample_ID)
```

### 9805 Dataset

Filter samples from `df_c` relevant to the 9805 study and compare with `df_9805`:

```{r backtrack-9805}
df_c_9805 <- df_c %>% 
  filter(!is.na(M_Recist_9805))

table(df_c_9805$Sample_ID %in% df_9805$Sample_ID)
